---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# respfun

This package is a collection of random functions for use with respirometry data. More will be added with time. Not intended to be a fully featured R package. 

### Installation
`respfun` can be installed using the `devtools` package:

```r
install.packages("devtools")
devtools::install_github("nicholascarey/respfun")
```

### Contents

For now there is just one function, `split_rate`. This function divides a metabolic rate between a group of individuals 
based on their masses and a metabolic scaling exponent. It was used in this recent publication:

Benjamin P. Burford, Nicholas Carey, William F. Gilly, Jeremy A. Goldbogen. 2019. Grouping reduces the metabolic demand of a social squid. 
*Marine Ecology Progress Series*. 612: 141–150 https://doi.org/10.3354/meps12880

I'll be revising the function soon to integrate nicely with [`respR`](https://github.com/januarharianto/respR) functions.


### Example

```{r echo = F}
library(respfun)
```

```{r eval = F}
## Simple example
split_rate(c(2, 3, 4, 5, 6), # body masses
           tR = 500,         # total metabolic rate of group
           b = 0.75)         # metabolic scaling exponent
```
Output:
```{r eval = T, echo = F}
## Simple example
split_rate(c(2, 3, 4, 5, 6), # body masses
           tR = 500,         # total metabolic rate of group
           b = 0.75)         # metabolic scaling exponent
```

### Future functionality

In due course I'll add a few more functions I find useful in working with respirometry and metabolic data.

For analysing your respirometry data check out my other package [`respR`](https://github.com/januarharianto/respR). 

