---
title: "`respfun`"
output:
  github_document:
    toc: true
    toc_depth: 4
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

[![Travis build status](https://travis-ci.org/nicholascarey/respfun.svg?branch=master)](https://travis-ci.org/nicholascarey/respfun)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/nicholascarey/respfun?branch=master&svg=true)](https://ci.appveyor.com/project/nicholascarey/respfun)
[![Coverage status](https://codecov.io/gh/nicholascarey/respfun/branch/master/graph/badge.svg)](https://codecov.io/github/nicholascarey/respfun?branch=master)

This package is a collection of functions for use with respirometry data and experiments. I will add to it periodically (suggestions [welcome](https://github.com/nicholascarey/respfun/issues)). It is not intended to be a fully featured R package, more a collection of handy functions. Similar functions are available in other packages and work perfectly well, I just like writing my own as a learning exercise. 

### Installation
`respfun` can be installed using the `devtools` package:

```r
install.packages("devtools")
devtools::install_github("nicholascarey/respfun")
```

### Functions

Currently there are six functions:

#### `split_rate`

This function divides a metabolic rate between a group of individuals based on their masses and a metabolic scaling exponent. It was used in this recent publication:

*Benjamin P. Burford, Nicholas Carey, William F. Gilly, Jeremy A. Goldbogen. 2019. Grouping reduces the metabolic demand of a social squid. 
**Marine Ecology Progress Series**. 612: 141–150 https://doi.org/10.3354/meps12880*

This function integrates with the [`respR`](https://github.com/januarharianto/respR) package: objects saved from the `respR::convert_rate` function can be 
entered (or `%>%` piped), and the rate and units will be automatically extracted. 

```{r echo = F}
library(respfun)
```

```{r eval = F}
## Simple example

split_rate(tR = 500,                  # total metabolic rate of group
           masses = c(2, 3, 4, 5, 6), # body masses
           b = 0.75,                  # metabolic scaling exponent
           units = "mg/h")            # units
```

```{r eval = T, echo = F}
## Simple example

split_rate(tR = 500,                  # total metabolic rate of group
           masses = c(2, 3, 4, 5, 6), # body masses
           b = 0.75,                  # metabolic scaling exponent
           units = "mg/h")            # units
```

```{r echo = F}
library(respR)
```

```{r eval = F}
## respR example

library(respR)                                           # Load respR

urchins.rd %>%                                           # NOT a group respirometry experiment,
                                                          # - Just using it as an example,
  inspect(1, 15) %>%                                     # inspect
  calc_rate(from = 4, to = 29, by = "time") %>%          # calculate rate
  print() %>%
  convert_rate(o2.unit = "mgl-1", time.unit = "m",       # convert
               output.unit = "mg/h", volume = 1.09) %>%
  split_rate(c(2, 3, 4, 5, 6, 7, 8),                     # divide
             0.75) %>%
  print()

```

```{r eval = T, echo = F, fig.keep = "none", warning = FALSE}
library(respR)

urchins.rd %>%                                           # NOT a group respirometry experiment,
                                                          # - Just using it as an example,
  inspect(1, 15) %>%                                     # inspect
  calc_rate(from = 4, to = 29, by = "time") %>%          # calculate rate
  print() %>%
  convert_rate(o2.unit = "mgl-1", time.unit = "m",       # convert
               output.unit = "mg/h", volume = 1.09) %>%
  print() %>%
  split_rate(c(2, 3, 4, 5, 6, 7, 8),                     # divide
             0.75) %>%
  print()
```

#### `eff_vol`
This calculates the 'effective volume' of a respirometer, that is the volume *minus* the volume of the specimen. This allows you to input the correct water volume into respirometry analyses in order to calculate the correct amount of oxygen used, for example in the [`respR::convert_rate`](https://januarharianto.github.io/respR/reference/convert_rate.html) function. 

The respirometer volume (`resp_vol`) can be corrected to get the effective volume in a number of ways:  

- A specimen volume (`spec_vol`) can be entered directly (if for instance you have calculated it geometrically or measured the displacement volume separately), in which case the effective volume is a simple subtraction.   

- Alternatively, you can enter the specimen mass and density (see `spec_density` function below), in which case the specimen volume is calculated and the correction performed using that. If your specimen is neutrally buoyant you could enter the density of the water here (seawater density is usually around 1026 kg/m^3, but see next option for precise calculation).

- Lastly, you can enter the specimen mass and make the assumption the specimen is the same density as the water, in which case temperature and salinity are required to calculate the water density and perform the correction (strictly speaking atmospheric pressure is also required; in reality it has a negligible effect within normal ranges, although the default value can be changed if desired). This is common in fish respirometry when the specimen is known to be neutrally buoyant, or nearly so. 

See `?eff_vol` for more. 

#### `spec_density`
This function calculates a specimen density from a total wet mass and a bouyant mass (see [here](https://www.researchgate.net/publication/266911357_Buoyant_weight_technique_Application_to_freshwater_bivalves)). This is common in studies of shell bearing species such as molluscs and echinoderms where it is a non-lethal way of examining the ratio of shell to tissue mass. It can also be used to determine the animal volume for correcting respirometry water volumes (see above). The water temperature and salinity at which buoyant mass was determined are required. 
See `?spec_density` for more. 

#### `wm_to_vol`
This is a simple function to convert a water mass to a water volume. Sometimes it is easier to weigh a respirometer to determine the water mass than try to measure the volume, for example across large size ranges of specimen and respirometer size (e.g. [Carey et al. 2016](https://www.dropbox.com/s/d4zp3vm6xakzkts/Carey%20et%20al%20JEB%202016.pdf?dl=0)), where systematic error is a concern. The water temperature and salinity are required.
See `?wm_to_vol` for more. 

#### `scale_rate`
This function scales a physiological rate to a different body mass using a scaling exponent. Works with both absolute (i.e. whole animal) and mass-specific rates. 
See `?wm_to_vol` for more. 

#### `q_ten`
Calculates any of the five parameters in the Q10 temperature relationship for physiological or chemical processes. Q10 describes the ratio by which a physiological or chemical rate changes with a 10°C increase in temperature. Essentially the same as the `Q10` function in the [`respirometry`](https://cran.r-project.org/web/packages/respirometry/index.html) package by Matthew Birk, although that has additional functionality for determining the best fit Q10 for a range of rates/temperatures. 
See `?q_ten` for more. 

### Future functionality
In due course I'll add a few more functions I find useful in working with respirometry and metabolic rate data. Suggestions for additional functions are welcome via [email](mailto:nicholascarey@gmail.com), or by [opening an issue](https://github.com/nicholascarey/respfun/issues).

### Full respirometry analyses
For analysing your respirometry data check out another package I co-developed: [`respR`](https://github.com/januarharianto/respR). 

